### 2022-01-13 paper/paper.tex
methodology: add new ui and more content

% Misc
\usepackage{gensymb}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
	\caption{Keyboard, camera angle, and placement to be used for testing.}
	\label{fig:metho-keeb}
	\centering
\end{figure}

@@

A computer vision algorithm will be created as a starting point for detecting,
and mapping the keyboard within a video. The flowchart of the algorithm is shown
in Figure~\ref{fig:metho-key-algo}.

@@

		\node[draw,
			below=of input,
			minimum width=3.5cm,
			minimum height=1cm] (bg-rem) {Background Removal};

		\node[draw,
			below=of bg-rem,
			minimum width=3.5cm,
			minimum height=1cm] (gray) {Convert to Grayscale};

@@

		\draw[-latex] (start) edge (input)
		(input) edge (bg-rem)
		(bg-rem) edge (gray)
		(gray) edge (sobel)
		(sobel) edge (contours)
		(contours) edge (map)
		(map) edge (output)
		(output) edge (end);

	\end{tikzpicture}
	\caption{Flowchart of Keyboard Detection and Mapping Algorithm}
	\label{fig:metho-key-algo}
\end{figure}

\subsubsection{Background Removal}
The background of the frame will be removed by using colors as the main metric
for determining the foreground and the background. This is made easier by the
setup as the setup will have a green background, with a light colored keyboard
as a foreground.

\subsubsection{Convert to Grayscale}
The frame will then be converted to grayscale. This is done because future steps
of the algorithm does not require color values and this improves the performance
of these algorithms.

\subsubsection{Edge Detection using Sobel Filter}
A Sobel Filter is a edge detection filter that uses a 3$times$3 kernel that is
convolved twice. Once horizontally, and another vertically to produce a
grayscale image of the outlines within the frame.

\subsubsection{Find Contours (Simple)}
Contours are curves joining all continuous points that have the same color or
intensity \parencite{opencv-contours}. In this algorithm, an OpenCV function is
to be used to find the contours of the outlines of the object found using the
previous step. This OpenCV function implements the algorithm of \cite{contours}.
The expected result is four contour points which are positioned on the four
edges of the keyboard.

\subsubsection{Overlay Virtual Keyboard Map using Contour Points}
The virtual keyboard map is a rectangular image that has individual region of
interest (ROI) for each key in a 60\% \ac{ansi} keyboard. Each key has a
corresponding color assigned to it and this color fills the region the key is
assigned to.

The virtual map is then stretched over the object, with the four contour points
as the four edges of the virtual keyboard map. The image generated is then
returned as the final output of the graph.

@@

A computer vision algorithm will be chosen as a starting point for detecting,
and tracking fingers within a video. The specific algorithm used for finger
detection and tracking will be the solution offered by MediaPipe dubbed
MediaPipe Hands. This algorithm is composed of multiple ML models working in
conjunction to be able to detect the different parts of the hands and track them
accurately.

\subsubsection{Palm Detection Model}
The first model, the Palm Detection Model detects the initial hand locations
using a single shot detector model based on the paper by \cite{ssd}. This
model achieves an average precision of 95.7\% in palm detection.

MediaPipe Hands detects the palms first, instead of the whole hands with one
model because the hands lack high contrast patterns. This prevents the model to
detect the hand accurately. In addition, detecting a palm is simpler compared to
detecting hands with articulated fingers since estimating a bounding box around
a rigid objects, i.e.\ a palm, is much simpler. Furthermore, a palm can be
modeled using only square anchors reducing the number of anchors by a factor of
3--5.

@@

This model was trained using 30,000 manually annotated, real-world images with
21 3D hand-knuckle coordinates. Using this information, the model can also
accurately add landmarks to partially visible hands and hands with
self-occlusion. This is also made possible by the model's consistent internal hand
pose representation.

@@

The second step is a continous loop where the mapping of the keyboard is used in
conjunction with the finger tracking algorithm chosen in
Section~\ref{section:metho-algo-finger}. The flowchart of the algorithm is shown
in Figure~\ref{metho-algo-integration}

@@

\subsubsection{Find ROI of the pressend key in Virtual Keyboard Map}
Since there is a known color-key mapping based on Virtual Keyboard Map, the
algorithm will find the correct color to search for based on the received keypress
packet. This color is isolated from the Virtual Keyboard Map, and the region of
interest and its coordinates are obtained.

\subsubsection{Find which fingertip is on that ROI}
The Finger Tracking Frame contains the pixel positions of each landmark of the
hand. For this step, the algorithm will find the landmark which is positioned
within the region of interest from the previous step.

\subsubsection{Return which fingertip}
The finger containing the landmark will be returned. This finger will then be
compared with the expected finger to be used to press the key to determine if it
is correct or not.

@@

\subsection{UX/UI design and Front-end development for the touch typing trainer}
The touch typing trainer will have its UX/UI design created in Figma. This
design will be based on other type tests such as Monkeytype
\parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
for showing real-time finger-key identification and mapping.
Figure~\ref{fig:ui-test} and \ref{fig:ui-stat} illustrates the two main pages of
the design.

>>>>

% Misc
\usepackage{gensymb}
\usepackage{amsmath}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
	\caption{Keyboard, camera angle, and placement to be used for testing.}
	\label{fig:metho-keeb}
\end{figure}

@@

A computer vision algorithm will be created as a starting point for detecting,
and mapping the keyboard within a video. The flowchart of the algorithm is shown
in Figure~\ref{fig:metho-algo-key-flow}.

@@

		\node[draw,
			below=of input,
			minimum width=3.5cm,
			minimum height=1cm] (gray) {Convert to Grayscale};

@@

		\draw[-latex] (start) edge (input)
		(input) edge (gray)
		(gray) edge (sobel)
		(sobel) edge (contours)
		(contours) edge (map)
		(map) edge (output)
		(output) edge (end);

	\end{tikzpicture}
	\caption{Flowchart of Keyboard Detection and Mapping Algorithm}
	\label{fig:metho-algo-key-flow}
\end{figure}

\subsubsection{Convert to Grayscale}
The frame will then be converted to a 256 level grayscale image. This step will
be performed because future steps of the algorithm will not require color values
to work. In addition, the performance of these steps will be improved.

\subsubsection{Edge Detection using Sobel Filter}
A Sobel Filter is an edge detection filter that uses a 3$\times$3 kernel that is
convolved twice. Once horizontally, and another vertically to produce a
grayscale image of the outlines within the frame. The kernels used by the Sobel
Filter \parencite{sobel2014} is shown in Figure~\ref{fig:metho-algo-key-sobel}.
This filter will be used in the algorithm to obtain the four edges of the
keyboard.

\begin{figure}[H]
	\centering
  $\begin{bmatrix}
  +1 & 0 & -1\\
  +2 & 0 & -2\\
  +1 & 0 & -1
  \end{bmatrix}$
	and
  $\begin{bmatrix}
  +1 & +2 & +1\\
  0 & 0 & 0\\
  -1 & -2 & -1
  \end{bmatrix}$
	\caption{Sobel Operator Kernels. Reproduced from \fullcite{sobel2014}}
	\label{fig:metho-algo-key-sobel}
\end{figure}

\subsubsection{Find Contours (Simple)}
\label{section:metho-algo-key-contours}
Contours are curves joining all continuous points that have the same color or
intensity \parencite{opencv-contours}. In this algorithm, an OpenCV function is
to be used to find the contours of the outlines of the object found using the
previous step. This OpenCV function implements the algorithm of \cite{contours}.
The expected result is four contour points which are positioned on the four
edges of the keyboard.

\subsubsection{Overlay Virtual Keyboard Map using Contour Points}
The virtual keyboard map is a rectangular image that contains individual region
of interest (ROI) for each key in a 60\% \ac{ansi} keyboard. Each key has a
corresponding color assigned to it and this color fills the region where this
key is located at. Figure~\ref{fig:metho-algo-key-map} is the virtual keyboard
map.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{image-map.png}
	\caption{Initial Image Map}
	\label{fig:metho-algo-key-map}
	\centering
\end{figure}

The virtual map will then be stretched over the object, with the four contour
points as the four edges of the virtual keyboard map. The image generated will
then be returned as the final output of the graph.
Figure~\ref{fig:metho-algo-key-overlay} illustrates this image, and
Figure~\ref{fig:metho-algo-key-half} shows how accurate the image map is when
overlaid over the image.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{image-map-final.png}
	\caption{Expected output of the algorithm}
	\label{fig:metho-algo-key-overlay}
	\centering
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{image-map-half.png}
	\caption{Overlaid image map shown at 60\% opacity}
	\label{fig:metho-algo-key-half}
	\centering
\end{figure}

@@

A computer vision algorithm will be chosen as a starting point for detecting,
and tracking fingers within a video. The specific algorithm used for finger
detection and tracking will be the solution offered by MediaPipe dubbed
MediaPipe Hands. This algorithm is composed of two ML models working in
conjunction to be able to detect the different parts of the hands and track them
accurately.

\subsubsection{Palm Detection Model}
The first model, the Palm Detection Model detects the initial hand locations
using a single shot detector model based on the paper by \cite{ssd}. This model
achieves an average precision of 95.7\% in palm detection
\parencite{mediapipe-hands}.

MediaPipe Hands detects the palms first, instead of the whole hands with one
model because the hands lack high contrast patterns. This prevents the model to
detect the hand accurately. In addition, detecting a palm is simpler compared to
detecting hands with articulated fingers since estimating a bounding box around
a rigid objects, i.e.\ a palm, is much simpler. Furthermore, a palm can be
modeled using only square anchors reducing the number of anchors by a factor of
3--5 \parencite{mediapipe-hands}.

@@

This model was trained using 30,000 manually annotated, real-world images with
21 3D hand-knuckle coordinates. Using this information, the model can also
accurately add landmarks to partially visible hands and hands with
self-occlusion. This is also made possible by the model's consistent internal
hand pose representation \parencite{mediapipe-hands}.

@@

The second step is a continuous loop where the mapping of the keyboard is used
in conjunction with the finger tracking algorithm chosen in
Section~\ref{section:metho-algo-finger}. The flowchart of the algorithm is shown
in Figure~\ref{fig:metho-algo-integration}

@@

\subsubsection{Find ROI of the pressend key in Virtual Keyboard Map}
Since there is a known color-key mapping based on the Virtual Keyboard Map, the
algorithm will find the correct color to search for based on the received
keypress packet. This color is isolated from the Virtual Keyboard Map, and the
region of interest and its coordinates are obtained. The edges of this ROI will
then be obtained by finding its contours. The specific implementation if this is
explained in Section~\ref{section:metho-algo-key-contours}.

The coordinate system will have its origin at the top left, starting at 0, 0.
Going to the right increases the value of the x-axis, and going to the bottom
increases the value of the y-axis. This will result in a coordinate system that
only has positive values with each value corresponding to a pixel.

The expected output of this is four sets of pairs. Each pair corresponds to an
edge of the region of interest.

\subsubsection{Find which fingertip is on that ROI}
The Finger Tracking Frame will contain the pixel positions of each landmark of
the hand. For this step, the algorithm will find the landmark which is
positioned within the region of interest obtained from the previous step. This
will be done using a series of checks:

\subsubsection{Return which fingertip}
The landmark will then be used to determine which specific finger corresponds to
the keypress. Figure~\ref{fig:metho-algo-integration-landmarks} shows each
possible landmark that may be returned from the previous step. This step will
then simply return the name of the landmark, up until the first underscore. As
an example, if the landmark found is \emph{MIDDLE\_FINGER\_DIP}, this step will
return \emph{MIDDLE} denoting that the middle finger is the finger that
corresponds with the keypress.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{hand-landmarks.png}
	\caption{Hand landmarks that may be returned by the algorithm. Reproduced from \fullcite{mediapipe-hands}}.
	\label{fig:metho-algo-integration-landmarks}
	\centering
\end{figure}

@@

\subsection{UX/UI design and Front-end development for the touch typing trainer}
The touch typing trainer will have its UX/UI design created in Figma. This
design will be based on other type tests such as Monkeytype
\parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
for showing real-time finger-key identification and mapping.
Figure~\ref{fig:ui-test} and \ref{fig:ui-stat} illustrates the two main pages of
the design. Additional screens, such as historical statistics, in-platform
tutorial, and a written guide are shown in the appendix.
